\renewcommand{\thesection}{S\arabic{section}}
\setcounter{section}{2}

\section{Epidemiological feature summaries for the combined illustrative example}
\label{sec:Appendix_S3}

Here we describe the epidemiological feature summaries in Fig.~\ref{fig:intro}G,H and explain why the seven-day growth-rate summary is especially sensitive to weakly identified high-frequency components.

\subsection{Epidemiological feature summaries}
\label{sec:appendix_feature_summaries}

We evaluated three quantities derived from each displayed upstream reference
trajectory: epidemic peak timing, epidemic peak magnitude, and the seven-day
log growth rate. All feature summaries were calculated over the displayed
interval from September 1, 2020, through March 31, 2021.

\subsubsection{Peak timing and peak magnitude}

For each displayed upstream reference trajectory, we first formed a reflected
14-day moving average,
\[
\bar U^{(14)}_{f_c}(t)
=
\frac{1}{14}
\sum_{s\in\mathcal W_t}
U^\star_{f_c}(s),
\]
where \(\mathcal W_t\) is the 14-day smoothing window used for the displayed trajectory, with reflection at the boundaries of the finite interval.
The peak index was defined as
\[
t^\star_{f_c}
=
\arg\max_t
\bar U^{(14)}_{f_c}(t).
\]

For visualization, relative peak timing was calculated as
\[
T^{\mathrm{rel}}_{f_c}
=
\frac{t^\star_{f_c}}
{T_{\mathrm{disp}}-1},
\]
where \(T_{\mathrm{disp}}\) is the number of days in the displayed analysis
interval and \(t^\star_{f_c}\) is indexed from zero within that interval.

The peak magnitude was evaluated from the unsmoothed cutoff-specific
reconstruction at the peak index selected from the smoothed trajectory,
\[
M_{f_c}
=
U^\star_{f_c}(t^\star_{f_c}).
\]
Relative peak magnitude was then defined as
\[
M^{\mathrm{rel}}_{f_c}
=
\frac{M_{f_c}}
{\max_{f_c'}M_{f_c'}}.
\]
The common 14-day smoothing used to locate the epidemic peak reduces the
sensitivity of the peak summaries to isolated daily fluctuations while
preserving the broad epidemic-wave structure.


\subsubsection{Seven-day log growth rate}
\label{sec:appendix_growth_rate}

For each displayed upstream reference trajectory, we calculated the seven-day
log growth rate
\[
r_{7,f_c}(t)
=
\frac{1}{7}
\log
\frac{
U^\star_{f_c}(t+7)+c
}{
U^\star_{f_c}(t)+c
},
\]
where \(c=1\) is a numerical offset used to avoid taking the logarithm of zero.
Values within the final seven days of the displayed interval do not have a
forward seven-day comparison and were filled from the nearest available growth-rate values before applying the common smoothing step. The resulting growth-rate trajectory was smoothed using a reflected five-day moving average. This final smoothing was applied identically to all displayed reference trajectories.


\subsection{Spectral sensitivity of the seven-day growth rate}
\label{sec:appendix_growth_spectral}

The seven-day growth rate is nonlinear in \(U(t)\), but its local spectral
sensitivity can be seen through a first-order perturbation. Let
\[
U_\epsilon(t)
=
U(t)+\epsilon\,\delta U(t).
\]
Differentiating \(r_7(t)\) with respect to \(\epsilon\) at \(\epsilon=0\)
gives
\[
\left.
\frac{\partial r_7[U_\epsilon](t)}
{\partial\epsilon}
\right|_{\epsilon=0}
=
\frac{1}{7}
\left[
\frac{\delta U(t+7)}
{U(t+7)+c}
-
\frac{\delta U(t)}
{U(t)+c}
\right].
\]

If the denominator varies slowly over a local interval, so that
\[
U(t+7)+c
\approx
U(t)+c
\approx
q(t),
\]
then
\[
\delta r_7(t)
\approx
\frac{
\delta U(t+7)-\delta U(t)
}{
7q(t)
}.
\]
The corresponding temporal-difference operator has frequency response
\[
H_7(f)
=
\frac{
e^{2\pi i 7f}-1
}{7},
\]
with magnitude
\[
|H_7(f)|
=
\frac{
2|\sin(7\pi f)|
}{7}.
\]
For low frequencies,
\[
|H_7(f)|
\approx
2\pi |f|,
\]
so the operator suppresses nearly constant components and increases its
sensitivity as frequency moves away from zero. Relative to summaries of the
overall level or broad epidemic-wave shape, the growth-rate calculation
therefore places greater emphasis on short-timescale differences in the
reconstructed trajectory.

The five-day moving average subsequently applied to the estimated growth rate
attenuates the highest-frequency numerical variation, but it does not remove
the underlying dependence of local growth estimates on upstream components
that are more weakly transmitted through the delay process.


Peak timing and peak magnitude are nonlinear global functionals and do not
possess a fixed linear transfer function analogous to \(H_7(f)\). Their
relative stability in Fig.~\ref{fig:intro}H is therefore an empirical
property of the admissible reconstruction family considered here, rather than
a universal theorem that peak summaries are always identifiable. The analysis
shows that, for this epidemic series and these reconstruction classes, the
broad peak structure is considerably more stable than the local growth-rate
trajectory.
