\renewcommand{\thesection}{S\arabic{section}}
\setcounter{section}{1}

\section{Construction of the combined illustrative example in Figure 1}
\label{sec:Appendix_S2}

Figure~\ref{fig:intro} provides a data-informed illustration of how delay filtering limits the temporal resolution of epidemic surveillance and how this limitation propagates to epidemiological features. The figure is intended to illustrate the observation and separability framework rather than to provide a substantive reconstruction of the U.S. epidemic.

\subsection{Observed downstream incidence}

Daily U.S. confirmed-case incidence was obtained from the Johns Hopkins University Center for Systems Science and Engineering COVID-19 data stream \cite{dong2020interactive}, accessed through the Delphi Epidata API \cite{farrow2015delphi}. We used the national-level \texttt{jhu-csse/confirmed\_incidence\_num} signal. The complete series from April 1, 2020, through June 30, 2021, was used to estimate the smoothed downstream mean, construct the candidate upstream trajectories, and perform convolution, thereby preserving the pre-window history required for observations near the beginning of the displayed interval. The spectral quantities, non-identifiable perturbation bands, and epidemiological feature summaries in Fig.~\ref{fig:intro} were evaluated over the displayed period from September 1, 2020, through March 31, 2021.

\subsection{Smoothed downstream mean and observation model}

To separate broad epidemic variation from day-to-day observational fluctuation, we estimated a smooth downstream mean \(\widehat{\mu}(t)\) using second-order trend filtering. The effective degrees of freedom were selected over a prespecified grid. Conditional on each candidate smooth mean, Gaussian, Poisson, and negative-binomial observation models were fitted, and the combination of trend-filter complexity and observation model was selected by the Akaike information criterion (AIC). The selected observation model was negative binomial for the displayed U.S. case series. Its fitted variance function was used to construct the local likelihood weights for the perturbation bands and the observation-variability reference shown in Fig.~\ref{fig:intro}F.

\subsection{Delay distribution}

The delay kernel was based on the published COVID-19 symptom-onset-to-case-report distribution reported by Zhang et al.~\cite{zhang2020evolving}. A lognormal representation was converted to a discrete daily probability mass function by integrating the fitted distribution over consecutive one-day intervals. The kernel was truncated at 60 days and renormalized to sum to one. The confirmed-case example used conversion rate \(p=1\).

\subsection{Frequency-constrained reference trajectories}

For each cutoff frequency \(f_c\), we represented the upstream trajectory in a real Fourier basis \(\Phi_{f_c}\) containing the constant component and sine and cosine components with frequencies satisfying \(0\leq f\leq f_c\). The cutoff-specific reference trajectory \(U^\star_{f_c}=\Phi_{f_c}\widehat\beta_{f_c}\) was obtained by solving
\[
\widehat\beta_{f_c}
=
\arg\min_{\beta:\,\Phi_{f_c}\beta\geq 0}
\left\{
\left\|\mathbf L_{\mathrm{fit}}\Phi_{f_c}\beta-\widehat{\mu}\right\|_2^2
+\lambda\|\beta_{-0}\|_2^2
\right\}
\]
where \(\mathbf L_{\mathrm{fit}}\) is the finite-window linear delay-observation matrix mapping upstream incidence to the downstream mean over the fitting window, \(\widehat{\mu}\) is the fitted smooth downstream mean, \(\beta_{-0}\) denotes all nonconstant Fourier coefficients, and \(\lambda=10^{-6}\) is a small ridge penalty used for numerical stability. Figure~\ref{fig:intro}A,C,G,H show cutoff frequencies \(f_c=1/60\), \(1/14\), and \(1/7\) cycles/day, corresponding to characteristic periods of 60, 14, and 7 days. We also constructed a no-cutoff reference by setting \(f_c=0.5\) cycles/day, the Nyquist frequency for daily data when present.

\subsection{Local non-identifiable perturbation bands}

The shaded bands in Fig.~\ref{fig:intro}A use the local likelihood approximation described in Appendix~\ref{sec:Appendix_S1}. For each cutoff \(f_c\), perturbations were restricted to the same Fourier subspace as the corresponding reference trajectory. Let \(\Phi_{f_c}\) denote this basis on the displayed window, \(\mathbf L_{\mathrm{disp}}\) the displayed-window linear delay-observation matrix, and \(\mathbf W_{\mathrm{loc}}\) the diagonal matrix of local likelihood weights under the selected observation model. The local downstream separability of a perturbation \(\delta U=\Phi_{f_c}b\) is
\[
J_{\mathrm{loc}}(\delta U)
=
\left\|\mathbf W_{\mathrm{loc}}^{1/2}\mathbf L_{\mathrm{disp}}\Phi_{f_c}b\right\|_2^2.
\]
Equivalently, writing \(\mathbf Z_{f_c}=\mathbf L_{\mathrm{disp}}\Phi_{f_c}\),
\[
J_{\mathrm{loc}}(\Phi_{f_c}b)
=
b^\top \mathbf I_{f_c}b,
\qquad
\mathbf I_{f_c}
=
\mathbf Z_{f_c}^\top \mathbf W_{\mathrm{loc}}\mathbf Z_{f_c}.
\]
For \(\alpha=0.20\), we used the non-identifiability threshold \(J_\alpha=4(1-\alpha)^2=2.56\), giving the perturbation ellipsoid
\[
\mathcal A_{f_c}
=
\left\{
b:
b^\top \mathbf I_{f_c}b
\leq
J_\alpha
\right\}.
\]
At each displayed time point \(t\), the pointwise perturbation radius was computed as
\[
B_{f_c}(t)
=
\sqrt{
J_\alpha\,
\phi_{f_c}(t)^\top
\mathbf I_{f_c}^{-}
\phi_{f_c}(t)
},
\]
where \(\phi_{f_c}(t)\) is the vector of Fourier-basis values at time \(t\), and \(\mathbf I_{f_c}^{-}\) is the Moore--Penrose inverse used in the numerical calculation, equal to the ordinary inverse when \(\mathbf I_{f_c}\) is full rank. The displayed band is
\[
\left[
\max\left\{
U^\star_{f_c}(t)-B_{f_c}(t),0
\right\},
\;
U^\star_{f_c}(t)+B_{f_c}(t)
\right].
\]
This band is a pointwise envelope of the local perturbation ellipsoid. The perturbation attaining the endpoint may differ across time points.

\subsection{Spectral display and observation-variability reference}

The spectral quantities in Fig.~\ref{fig:intro}D--F were computed over the displayed period using the same one-sided discrete Fourier transform and frequency grid. Tilde notation denotes the corresponding discrete Fourier coefficient on this displayed window; for example, \(\Delta\widetilde U_{f_c,m}\) is the coefficient of \(\Delta U_{f_c}\) at frequency \(f_m\). These panels display how upstream spectral contrast is attenuated before it appears in the downstream mean; separability itself is evaluated using the likelihood-weighted quadratic form described in Methods. Panel D shows, for each cutoff-specific reference trajectory, the upstream spectral contrast relative to the no-cutoff reference,
\[
\begin{aligned}
P_{\Delta U,f_c}(f_m)
&=
\frac{q_m}{T_w}
\left|
\Delta\widetilde U_{f_c,m}
\right|^2,\\
\Delta U_{f_c}
&=
U^\star_{f_c}-U^\star_{\mathrm{no\ cutoff}}.
\end{aligned}
\]
where \(T_w\) is the displayed-window length and \(q_m\) is the standard one-sided multiplicity weight. In the displayed spectra, the zero-frequency component is omitted because it represents mean-level contrast rather than temporal variation. For the one-sided display, \(q_m=2\) for positive interior frequencies and \(q_m=1\) for boundary frequencies, including the zero-frequency component and the Nyquist frequency \(f=0.5\) cycles/day when it is present. Panel E shows the squared delay attenuation \(|\widetilde G_m|^2\), where \(\widetilde G_m\) is the discrete Fourier coefficient of the delay kernel at frequency \(f_m\). Panel F shows the corresponding downstream-mean spectral contrast computed from the displayed portions of the full delay-convolved reference means,
\[
\begin{aligned}
P_{\Delta m,f_c}(f_m)
&=
\frac{q_m}{T_w}
\left|
\Delta\widetilde m_{f_c,m}
\right|^2,\\
\Delta m_{f_c}
&=
\Pi_{\mathrm{disp}}\mathbf L_{\mathrm{fit}} U^\star_{f_c}
-
\Pi_{\mathrm{disp}}\mathbf L_{\mathrm{fit}} U^\star_{\mathrm{no\ cutoff}}.
\end{aligned}
\]
Here \(\Pi_{\mathrm{disp}}\) denotes restriction to the displayed period, so the downstream spectra are computed after full-window convolution and then restricted to the displayed period.

The horizontal line in Fig.~\ref{fig:intro}F is a visual reference for the scale of observation variability on the one-sided frequency scale. For Poisson and negative-binomial observations, we compute local likelihood weights \(w_t\) at the fitted downstream mean and display
\[
2\,\operatorname{median}_{t\in \mathrm{disp}}\{1/w_t\},
\]
where the median is taken over the displayed period. The factor of two accounts for the pairing of positive- and negative-frequency coefficients at interior frequencies in the one-sided display. Separability calculations use the likelihood-weighted quadratic form described in Methods.
